A Bayesian ranking statistic to find high-mass black holes in LIGO--Virgo data

### 2021-03-10 main.tex

\usepackage{aas_macros}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{array}
\usepackage{color,units}
\usepackage[dvipsnames]{xcolor} % for more colours!
\usepackage{lineno}
\usepackage{dcolumn}
\usepackage{longtable}
\usepackage[normalem]{ulem} %% for striking out text
\usepackage{subfigure}
\usepackage[T1]{fontenc}
\usepackage[breaklinks]{hyperref}
\usepackage{booktabs}
\usepackage{xspace} 
\usepackage{bm}% bold math
\usepackage{graphicx} % Include figure files
\graphicspath{{images/}} %Setting the graphicspath

@@

\title{A Bayesian ranking statistic to find \\high-mass black holes in LIGO data}

% \author{Author list TBD}
\input{authors}

@@

\begin{abstract}
The detection of intermediate-mass black holes ( $10^2-10^6$~\msun) will shed light on the formation of supermassive black holes and thus galaxy formation. Although LIGO is sensitive to the merger of binary black holes with total masses up to $400$~\msun, only 4 of their 50 detections have a total mass $>100$~\msun. A possible explanation for the absence of intermediate-mass events may be their misclassification as short-duration instrumental noise transients. Short-duration instrumental transients mimic the short-duration gravitational-wave signals from intermediate-mass binary black hole mergers. Here we demonstrate that a ranking statistic utilising Bayesian inference could be a more sensitive tool for detecting high-mass binary black hole mergers (systems with a total mass $>55$~\msun) compared to a traditional match-filter ranking statistic. We applied this technique on the high-mass triggers during LIGO's second observing run to investigate the possibility of discovering new gravitational-wave signals from high mass black hole binaries and re-calculate the significance of high-mass candidate events. Although our analysis does not discover new gravitational wave events, it does provide a Bayesian measure of various candidates and events. 
\end{abstract}

@@

%%%%%%---SECTIONS---%%%%%%%%%%%%
\section{\label{sec:Introduction}Introduction}

Since the 1970s, there has been an accumulation of evidence for stellar-mass and supermassive black holes. In April 2019, the Event Horizon Telescope provided the first visual evidence of the supermassive black hole M87~\cite{m87photo}. As of January 2021, the LIGO Scientific Collaboration has confirmed $\sim50$ binary black hole systems and listed numerous candidate events~\cite{GWTC1, GWTC2, gracedb}. Additionally, since the public release of LIGO's first and second observing run's data (O1 and O2), several external groups, such as \pycbc~\cite{pycbc_code} and a research team at the Institute for Advanced Study (IAS)~\cite{IAS0, IAS1, IAS2}, have independently searched and identified additional binary black hole gravitational wave candidates. These various discoveries have firmly authenticated the existence of stellar-mass black holes, supermassive black holes and binary black hole systems.

Until recently, there was no definitive evidence for intermediate-mass black holes, the black holes whose masses lie between that of stellar-mass and supermassive black hole systems ($10^2-10^6\ \msun$). This changed with the detection of GW190521, a unique gravitational wave event that led to the formation of a black hole with a mass  $142\ \msun$, the first confirmed direct discovery of an intermediate-mass black hole~\cite{gw190521}. Although this is the first gravitational wave event leading to discovering a black hole with a mass in the intermediate range, ground-based gravitational wave detectors are sensitive to gravitational waves from even more massive systems to systems with a total mass of $400\ \msun$~\cite{GWTC1}. Additionally, gravitational waves from systems with masses greater than $100\ \msun$ systems should occur at a rate of $0-10\text{yr}^{-1}$~\cite{fregeau2006imbhbRatePrediction, mandel2008rates,rodriguez2015bbhRatePredictions,ligo_imbh_search}.  \av{this is surely outdated... get more recent rate quote} Hence, LIGO might have gravitational wave events from more intermediate-mass systems hidden in its data, potentially even in O2 data.

However, even after conducting a targeted matched-filter based search for gravitational waves from intermediate-mass black holes in O2's data, the heaviest total mass detected so far is approximately $80\ \msun$~\cite{GWTC1,imbhbSearch2014, ligo_imbh_search}. A possible explanation for the absence of intermediate-mass events may be due to their misclassification as short-duration instrumental noise transients known as glitches~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. These glitches can mimic astrophysical signals and hence decrease the significance of true gravitational wave events.

One method to account for glitches while searching for gravitational waves from coalescing compact binaries is by utilising an astrophysical Bayesian odds~\cite{bci, BCR1, BCR2, bcr_gw151216, bayesian_odds}. A true Bayesian odds calculated without using bootstrap techniques can provide events with a more accurate significance that is agnostic to a specific search strategy~\cite{BCR2, bcr_gw151216,  bayesian_odds}.

Additionally, Bayesian odds can include comprehensive astrophysical information in its calculation. For example, a Bayesian odds to help detect gravitational-wave candidates can incorporate information on if an event's binary system was precessing and if the signal contains higher-order modes and is coherent amongst various detectors. Because of the possibility of Bayesian methods to incorporate detailed physical information about a gravitational wave signal, the LIGO Scientific Collaboration uses Bayesian methods to determine the source parameters of gravitational-wave events~\cite{abbott2016ligo, GWTC1}. This paper demonstrates that the power of Bayesian methods used in parameter estimation can also discriminate between coherent gravitational-wave signals, incoherent glitches, and Gaussian noise in the form of a Bayesian odds, utilised as a ranking statistic.

In this paper, we utilise a Bayesian method, called the Bayesian Coherence Ratio \bcr~\cite{BCR1}, to rank the candidate gravitational wave signals from high-mass compact binary coalescences (systems with total masses in the range of $55-500\ \msun$) in the detector data recorded during O2. Although the \bcr, utilising bootstrap techniques, does not provide the true Bayesian odds, it provides a more astrophysical measure of candidate events' significance than a traditional match-filter significance.

We find that (a) high-mass events reported in the GWTC-1, including GW170729 (an event with disputed \pastro amongst various search pipelines) have high significance; (b) high-mass events detected from the IAS group have differing levels of significance; and that (c) our ranking statistic does not identify any intermediate-mass black holes, but does identify an unreported stellar-mass binary black hole candidate, 170222.

The remainder of this paper is structured as follows. We outline our methods, including details of the \bcr and the retrieval of our candidate events in Section~\ref{sec:method}. We present details on the implementation of our analysis in Section~\ref{sec:Analysis}. Finally, we present our results in Section~\ref{sec:Results}, and discuss these results in the context of the significance of gravitational wave candidates in Section~\ref{sec:Conclusion}.

\section{Method\label{sec:method}}
The gravitational wave community uses Bayesian inference to perform parameter estimation and model selection. In this work, we utilise Bayesian inference to calculate the significance of high-mass candidate events in O2 by using the \bcr as a ranking statistic, taking a step toward building a unified Bayesian framework to search for candidates and estimate their parameters.

Although a dedicated Bayesian search for gravitational waves, as presented by ~\citet{BCR2}, does not require noise estimation using empirical methods, this Bayesian significance ranking technique utilises \textit{time-slides}. To perform time-slides, data from independent observatories are time-shifted by amounts greater than the light-travel time between the two detectors. Each unique time-slide amount creates an artificial signal-free `background' data set. When search pipelines scan these background data sets for gravitational wave events, they can find \textit{triggers}\footnotemark, even though the background data set should not contain coherent astrophysical signals. These triggers in background data are labelled background triggers, while coherent triggers obtained in non-time-slid data are labelled candidate triggers.

@@

@@

The \bcr is a Bayesian odds ratio like the above, of a coherent signal hypotheses $\mathcal{H}_S$ and an incoherent instrumental feature hypothesis $\mathcal{H}_I$ for a network of $D$ detectors. $\mathcal{H_I}$ states that each detector $i$ has either pure Gaussian noise $\mathcal{H}_N$ or a glitch $\mathcal{H}_G$. Following~\citet{BCR1}, the \bcr is given by
\begin{equation}
\label{eq:bcr}
\bcr = \frac{\alpha Z^S}{\prod\limits^D_{i=1} \ [\beta Z^G_i + (1-\beta)Z^N_i]}\ ,
\end{equation}
where $Z^S$, $Z^G_i$ and $Z^N_i$ are the Bayesian evidences (marginalised likelihoods) for $\mathcal{H}_S$, $\mathcal{H}_N$, and $\mathcal{H}_G$. $\alpha$ and $\beta$, are the prior odds for obtaining a signal $\alpha=P(\mathcal{H}_S)/P(\mathcal{H}_I)$ and the prior odds for obtaining a glitch $\beta=P(\mathcal{H}_G)/P(\mathcal{H}_I)$. As the rate of signal and glitches are unknown, these priors $\alpha$ and $\beta$ are tuned to maximise the \bcr distributions for background data (coherent signal-free data) and simulated signals~\cite{BCR1}.

\subsection{Bayesian Evidence Evaluation}
\subsubsection{Noise Model}
We assume that each detector's noise is Gaussian and stationary over the period being analysed~\cite{ligo_psd}. In practice, we assume that the noise has a mean of zero that the noise variance $\sigma^2$ is proportional to the noise power spectral density (PSD) \psd of the data. Using the \psd, for each data segment $d_i$ in each of the $i$ detectors in a network of $D$ detectors, we can write 
\begin{equation}
\label{eq:zn}
Z^N_i = \mathcal{N}(d_i) = \frac{1}{2\pi \psd_i} \ \text{exp}\left(-\frac{1}{2} \frac{d_i}{\psd_i} \right) \ ,
\end{equation}
where $\mathcal{N}(d_i)$ is a normal distribution with $\mu=0$ and $\sigma^2\sim \psd$.

@@

>>>>

\usepackage{aas_macros}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{array}
\usepackage{color,units}
\usepackage[dvipsnames]{xcolor} % for more colours!
\usepackage{lineno}
\usepackage{dcolumn}
\usepackage{longtable}
\usepackage[normalem]{ulem} %% for striking out text
\usepackage{subfigure}
\usepackage[T1]{fontenc}
\usepackage[breaklinks]{hyperref}
\usepackage{booktabs}
\usepackage{xspace} 
\usepackage{bm}% bold math
\usepackage{graphicx} % Include figure files
% \usepackage{authblk}

\graphicspath{{images/}} %Setting the graphicspath

@@

\title{A Bayesian ranking statistic to find high-mass black holes in LIGO--Virgo data}

% \input{authors}
%\author[1]{Test}
%\affil[1]{Test Affil}

@@

\begin{abstract}
The detection of intermediate-mass black holes ( $10^2-10^6$~\msun) will shed light on the formation of supermassive black holes and thus galaxy formation. Although LIGO is sensitive to the merger of binary black holes with total masses up to $400$~\msun, only 4 of their 50 detections have a total mass $>100$~\msun with >95\% credibility. A possible explanation for the absence of intermediate-mass events may be their misclassification as short-duration instrumental noise transients. Short-duration instrumental transients mimic the short-duration gravitational-wave signals from intermediate-mass binary black hole mergers. 
\et{I think the preceding text is misleading. We make it sound like we expect a uniform distribution of total mass, and so it's a surprising that we don't see IMBH. However, in reality, it is theoretically challenging to make IMBH. You need an exceptionally massive progenitor star to avoid pair instability supernova. Or, maybe you can make IMBH from hierarchical mergers or accretion. No one expects them to be common, but they are  interesting because of their connection to supermassive black holes. I suggest rewriting the preceding text to explain how IMBH might form and why they are interesting to look for.}
Here we demonstrate a ranking statistic utilising Bayesian inference for the detection of high-mass binary black hole mergers (with a total mass $>55$~\msun). \et{Claims of beating matched filtering not yet borne out.} We apply this technique on the high-mass triggers during LIGO's second observing run to search for previously unresolved gravitational-wave signals from high-mass binary black holes. Our analysis does not discover new gravitational-wave events. However, we find support for some previously identified borderline detections.
\et{Don't oversell Bayesian measure since our method is only semi-Bayesian.}
\end{abstract}

@@

%%%%%%---SECTIONS---%%%%%%%%%%%%
\section{New Intro}
In the 1960s, X-ray and optical observations from the Cyg-X1 binary star system led to the discovery of the first stellar-mass black hole~\cite{}. Shortly afterwards, in 1974, kinematic measurements of stars near the center of our galaxy demonstrated the presence of a supermassive black hole of mass $10^6\ \text{M}_{\odot}$ -- Sag A$^*$~\cite{}. These two detections firmly established the existence of stellar-mass
black holes $\text{M}_\text{BH} < 10^{2} \ \text{M}_{\odot}$ and
supermassive black holes, $\text{M}_\text{BH} > 10^{5} \ \text{M}_{\odot}$ in our universe. Since then, astronomers accumulated more definitive evidence for stellar-mass and supermassive black holes. However, a gap in between the mass-ranges of these two categories began to form: there is a deficiency of black holes in the intermediate-mass range $10^{2} - 10^{5}\ \text{M}_{\odot}$. These intermediate-mass black holes must have existed to form the supermassive black holes detected to exist in the early universe~\cite{}. However, we only have a handful of candidates and one confirmed intermediate-mass black hole~\cite{}. The study and discovery of intermediate-mass black holes will help bridge the gap and provide a pathway to forming supermassive black holes.

Intermediate-mass black holes may form various scenarios ranging from hierarchical mergers of stellar-mass black holes to large metal-free stars' direct collapse. Hierarchical mergers in active galactic nuclei migration-traps, globular and stellar clusters may be a feasible method for black holes to accumulate mass and reach the intermediate-mass range. However, when black holes merge, they can receive recoil-kicks up to $\sim$2,500 km/s. Hence, the merger remnants require their recoil kick to be less than its environment's escape velocity to undergo successive mergers. This tight constraint makes hierarchical mergers a challenging scenario.

Another pathway to form intermediate-mass black involves the direct collapse of massive, metal-free Population III stars. Typically, massive stars $> 100 \ \msun$ can eject more than half of their mass in the form of radiatively driven stellar winds. However, at zero metallicity, mass loss through stellar winds is negligible (Kudritzki 2000). Hence,  Population III stars have the opportunity to complete their life-cycles without losing a significant portion of their mass. Moreover, Population III stars with masses above $230 \ \msun$ will avoid electron-positron pair-instability. Consequently, oxygen and silicon burning will not drive an explosion, allowing the star to undergo a complete collapse to leave behind an intermediate-mass black hole remnant with a mass of $120-280 \ \msun$, dependant on the initial star mass (Bond et al. 1984). The intermediate-mass black holes that form as remnants of these large Population III stars may exist today in clusters located at the galaxies' bulges. If these bulges contain intermediate-mass black holes, they could also provide evidence for Population III stars' existence.

As there is inadequate observational evidence for intermediate-mass black holes, it is challenging to state which formation path is favoured. However, the observational work does allude to the existence of intermediate-mass black holes. The search for intermediate-mass black holes is underway using direct kinematic measurements, reverberation mapping, studying luminosity from low-luminous active galactic nuclei and Ultraluminous X-ray Sources.

Dynamic measurements of individual sources orbiting an intermediate-mass black hole is that the sphere of influence too small to resolve motions. Can only be used on close sources. The issue with reverberation mapping and the other EM spectra, cant be sure of anisotropic emission, non-linear emission. large uncertainties. \url{https://iopscience.iop.org/article/10.3847/2041-8213/ab745b}

The most promising avenues to find $100−1000\ \msun$ black holes are either to identify stellar binaries, which are expected to be common or to detect them in gravitational radiation. Ground-based gravitational wave limits on the low end of this mass regime are continuing to improve.

LIGO found 190521, can find more if we are more sensitive, and if we can better harness data

\section{\label{sec:Introduction}Introduction}

Since the 1970s, there has been an accumulation of evidence for stellar-mass and supermassive black holes. In April 2019, the Event Horizon Telescope provided the first visual evidence of the supermassive black hole M87~\cite{m87photo}. As of January 2021, the LIGO Scientific Collaboration has confirmed $\sim50$ binary black hole systems and listed numerous candidate events~\cite{GWTC1, GWTC2, gracedb}. Additionally, since the public release of LIGO's first and second observing run's data (O1 and O2), search pipelines such as \pycbc~\cite{pycbc_code} and a research team at the Institute for Advanced Study (IAS)~\cite{IAS0, IAS1, IAS2}, have independently identified additional binary black hole gravitational wave candidates.

\et{Add a paragraph here discussing basic ideas in IMBH theory. Explain why the are not easy to make. Discuss scenarios by which they can be made. Discuss the role they may play in the formation of supermassive BH. Then, we transition to the next paragraph about observational evidence.}

Until recently, there was no definitive evidence for intermediate-mass black holes with a mass in between the range of stellar-mass and supermassive black holes ($10^2-10^6\ \msun$). 
\et{There have been numerous claims of IMBH detection. The previous sentence is correct that there are no \textit{definitive} detections before GW190521, but the paragraph is misleading by omission. Insert here an exhaustive review of previous IMBH detection claims, noting who made the claim, the mass, and the method.}
This changed with the detection of GW190521, a unique gravitational wave event that led to the formation of a black hole with a mass  $142\ \msun$, the first confirmed direct discovery of an intermediate-mass black hole~\cite{gw190521}. Although this is the first gravitational-wave observation of an IMBH, ground-based gravitational wave detectors are sensitive to gravitational waves from even more massive systems with a total mass of up to $400\ \msun$~\cite{GWTC1}.

\et{The following paragraph tries to solve a problem that doesn't exist: as noted in comments above, the absence of GWs from IMBH is not surprising. Delete or rewrite if you want to emphasise the difficulty in IMBH detection.}
A possible explanation for the absence of intermediate-mass events may be due to their misclassification as short-duration instrumental noise transients known as glitches~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. These glitches can mimic astrophysical signals and hence decrease the significance of true gravitational wave events.

One method to account for glitches while searching for gravitational waves from coalescing compact binaries is by utilising the astrophysical Bayesian odds~\cite{bci, BCR1, BCR2, bcr_gw151216, bayesian_odds}. A true Bayesian odds, calculated without using bootstrap techniques, can provide more accurate estimate of significance that does note depend on the search pipeline~\cite{BCR2, bcr_gw151216,  bayesian_odds}.
In this paper, we utilise a Bayesian method, called the Bayesian Coherence Ratio \bcr~\cite{BCR1}, to rank the candidate gravitational wave signals from high-mass compact binary coalescences (systems with total masses in the range of $55-500\ \msun$) in the detector data recorded during O2. Although the \bcr, utilising bootstrap techniques, does not provide the true Bayesian odds, it provides a more astrophysical measure of candidate events' significance than a traditional match-filter significance.
\et{This pitch, that the method is ``a more astrophysical measure'' is unconvincing. (What does that even mean?) Chat with RS/ET to come up with a better pitch built around the idea that the detection statistic is built on Bayesian inferences results, and that it's a step toward fully Bayesian detection.}

We find that (a) high-mass events reported in the GWTC-1, including GW170729 (the least significant event in GWTC-1) have high significance; (b) high-mass events detected from the IAS group have differing levels of significance; \et{differing from what? edit this passage to state more clearly that we find statistical support for some of their candidates, but not for others} and that (c) our ranking statistic does not identify any intermediate-mass black holes, but does identify an unreported stellar-mass binary black hole candidate, 170222 \et{State statistical significance.}.

The remainder of this paper is structured as follows. We outline our methods, including details of our ranking statistic and the retrieval of our candidate events in Section~\ref{sec:method}. We present details on the implementation of our analysis in Section~\ref{sec:Analysis}. Finally, we present our results in Section~\ref{sec:Results}, and discuss these results in the context of the significance of gravitational-wave candidates in Section~\ref{sec:Conclusion}.

\subsection{Notes for motivation}
\begin{itemize}
    \item if truly Bayesian -- then its more astrophysical 
    \item we're focused on odds -- but using it as a ranking statistic
    \item GOAL: look for an IMBH, not to perfect the statistic     
    \item  detection statistic that is based on a Bayesian approach -- a different way of looking for IMBHs, in place of matched-filter
    \item normal statistic -- is maximum likelihood appraach (highest SNR), this one marginalises over all the templates -- closer to an "odds" like goal
    \item this is NOT BCR2, this is a ranking statistic 
    \item between Isi et al and Ashton et al,  more like Ashton and not matched filtering 
    \item we didnt do a full Ashton because (no need to argue why we didnt do this)
    \item whats out there, where does this fit into lit, what are we doing vs not doing 
    \item  reproduced all these other results 
    \item new in this paper: using Bayesian coherence (an ingredient in Ashton et al) and looking for NEW signals in triggers, the PRatten et al looks at all the catalogued triggers. 
\end{itemize}

\begin{itemize}
    \item  IMBH signals look a lot like glitches
    \item  We compute \bcr which needs $\alpha$ $\beta$
    \item  We don't know the true probability $P(\mathcal{H}_S), P(\mathcal{H}_I)$ 
    \item hence we cannot take \bcr at face value because we dont know the true distributions of the above
    \item Studying p(\bcr| d=background) p(\bcr| d=signals) gives us confidence to discriminate between signals and glitches
    \item  we "rank" the BCR of foreground triggers based on that of background triggers which allows us to compute a frequentist estimator for its significance.
\end{itemize}

\section{New Method\label{sec:method}}

The gravitational wave community uses Bayesian inference to perform parameter estimation and model selection. This work utilises Bayesian inference to calculate the significance of high-mass candidate triggers\footnote{\textit{Trigger:} point in time when the matched-filter SNR is larger than a threshold value for a given gravitational wave template~\cite{pycbc_og1}. \textit{Candidate Trigger:} trigger that may be from astrophysical origins.} in LIGO's second observing run. We calculate the significance of triggers using the Bayesian Coherence Ratio~\cite{BCR1}, \bcr, as a ranking statistic, a step toward building a unified Bayesian framework to search for candidates and estimate their parameters.

The \bcr is a Bayesian odds ratio like the above, of a coherent signal hypotheses $\mathcal{H}_S$ and an incoherent instrumental feature hypothesis $\mathcal{H}_I$ for a network of $D$ detectors. $\mathcal{H_I}$ states that each detector $i$ has either pure stationary Gaussian noise $\mathcal{H}_N$ or Gaussian noise and an incoherent noise transient (glitch). Taking $\mathcal{H}_G$, $Z^S$, $Z^G_i$ and $Z^N_i$ as the Bayesian evidences (marginalised likelihoods) for $\mathcal{H}_S$, $\mathcal{H}_N$, and $\mathcal{H}_G$, the \bcr is given by
\begin{equation}
\label{eq:bcr}
\bcr = \frac{\alpha Z^S}{\prod\limits^D_{i=1} \ [\beta Z^G_i + (1-\beta)Z^N_i]}\ ,
\end{equation}
where $\alpha$ and $\beta$ are the prior-odds of obtaining a signal or a glitch from a stretch of data. The prior-odds can be defined more explicitly as 
\begin{itemize}
    \item $\alpha=P(\mathcal{H}_S)/P(\mathcal{H}_I)$, the prior-odds for obtaining a coherent signal versus an incoherent instrumental feature.
    \item $\beta=P(\mathcal{H}_G| \mathcal{H}_I)$, the prior odds for obtaining a glitch assuming there is an incoherent instrumental feature.
\end{itemize}

When $\mathcal{H}_S$ and $\mathcal{H}_I$ are precisely described and the correct prior-odds are known, the \bcr is a Bayesian odds-ratio. As an odds-ratio, the \bcr is the optimal discriminator between coherent signals and incoherent instrumental features. However, as the priors-odds are unknown, it is invalid to use the \bcr as an odds-ratio. Instead of interpreting the \bcr as a Bayesian odds-ratio, it can be used as a ranking statistic. Using the \bcr as a ranking statistic we can obtain a frequentist significance of a candidate trigger \bcr measured against a {background trigger}\footnote{\textit{Background Triggers}: triggers that do not contain a coherent gravitational wave signal.} \bcr distribution.

When using the \bcr as a detection statistic, the physical interpretation of the prior-odds is lost. Hence, the prior-odds are empirically tuned to maximise the separation between the \bcr distributions for background triggers (expected to favour the $\mathcal{H}_I$ hypothesis) and {simulated triggers}\footnote{\textit{Simulated Triggers}: artificial triggers manufactured by injecting signals into the detector data.} (expected to favour the $\mathcal{H}_S$ hypothesis). Increasing the separation between the two distributions can improve ability of the \bcr to discriminate candidate events as coherent signals or incoherent instrumental features. The tuning process is described in detail in Appendix~\ref{sec:tuning-prior-odds}.

\subsection{Calculating the Significance of Candidate events}
Candidate trigger \bcr values are either statistically insignificant compared to the background \bcr distribution, implying the candidate is more probable to be an incoherent instrumental feature (the $\mathcal{H}_I$ hypothesis), or statistically significant to the background distribution, indicating the possible presence of an astrophysical signal (the $\mathcal{H}_S$ hypothesis). A false alarm probability with trial factors \fap for the candidate triggers can quantify the significance. The \fap is the probability that a candidate trigger originating from a non-astrophysical source can be is falsely identified as a signal.

To calculate the \fap, each candidate \bcr is considered as a single statistical trail that can occur at a fixed false alarm probability $f$, where $f$ is the probability of observing a background \bcr greater than or equal to the candidate \bcr,
\begin{equation}
    f = \frac{\text{Count of } \bcr' \leq \bcr}{\text{Count of } \bcr'} \ .
\end{equation}
The false alarm probability with trails \fap that the \bcr measurement occurs at least once for $N$ trials $(N > 0)$, where $N$ is the number of candidate triggers is
\begin{equation}
    \fap = 1 - (1-f)^N \ .
\end{equation}

Finally, the \fap can be used to construct a \pastro, the probability that a signal is of astrophysical origin~\cite{pastro_1,pastro_2,pastro_3}
\begin{equation}
    \pastro = 1 -  \fap \ .
\end{equation}

For \pycbc to consider a trigger to be a candidate trigger, a trigger that may be from astrophysical origins, the trigger must be observed between detectors with the same template and a time of arrival difference less than the gravitational-wave travel time~\cite{pycbc_og1}. To test its pipeline's sensitivity, \pycbc also conducts searches for simulated triggers. Finally, to quantify the statistical significance of candidate events, \pycbc artificially constructs background triggers to compare against the candidate events. These background triggers are coherent signal-free events, constructed by applying relative offsets, or time-slides, between the data of different detectors~\cite{pycbc_og6}. The time-slides to generate the background triggers are greater than the gravitational-wave travel time between detectors to ensure that the background triggers are not of astrophysical origins.

\section{Method\label{sec:method}}

\rs{I like the idea of the general intro: I suggest saying that you'll begin by introducing the Bayesian formalism, then defning the BCR, then discussing how we construct a ranking statistic using foreground and background. As it stands, there's a lot of discussion about how the BCR differs to other searches, and some technical details, but you haven't yet said what it *is*.
Before talking about how the BCR differs from other searches etc.., I think there needs to be a description of what it is and where it comes from. (essentially describing everything up to the current eq. 4). This sets up the framework, and you can then talk about why we construct it for both foreground and background to get a ranking statistic. }.
The gravitational wave community uses Bayesian inference to perform parameter estimation and model selection. In this work, we utilise Bayesian inference to calculate the significance of high-mass candidate events in O2 by using the \bcr as a ranking statistic, taking a step toward building a unified Bayesian framework to search for candidates and estimate their parameters.

\rs{Suggest revisiting the rest of this intro to section 2 after moving the discussion of Bayes theorem and the BCR.}
Although a dedicated Bayesian search for gravitational waves, as presented by ~\citet{BCR2}, does not require noise estimation using empirical methods, this Bayesian significance ranking technique utilises \textit{time-slides}. To perform time-slides, data from independent observatories are time-shifted by amounts greater than the light-travel time between the two detectors. Each unique time-slide amount creates an artificial signal-free `background' data set. When search pipelines scan these background data sets for gravitational wave events, they can find \textit{triggers}\footnotemark, even though the background data set should not contain coherent astrophysical signals. These triggers in background data are labelled background triggers, while coherent triggers obtained in non-time-slid data are labelled candidate triggers.

@@

\subsection{Triggers for Analysis}
\rs{I suggest moving this section to after you define the BCR and the need to estimate its significance empirically}
The LIGO Scientific collaboration operates several search pipelines that scan for gravitational waves from compact binary mergers such as \texttt{GstLAL}, \texttt{MBTA}, \texttt{SPIIR} and \pycbc~\cite{GWTC1}.

@@

The \bcr is a Bayesian odds ratio like the above, of a coherent signal hypotheses $\mathcal{H}_S$ and an incoherent instrumental feature hypothesis $\mathcal{H}_I$ for a network of $D$ detectors. $\mathcal{H_I}$ states that each detector $i$ has either pure stationary Gaussian noise $\mathcal{H}_N$ or an incoherent noise transient (glitch) $\mathcal{H}_G$. Following~\citet{BCR1}, the \bcr is given by
\begin{equation}
\label{eq:bcr}
\bcr = \frac{\alpha Z^S}{\prod\limits^D_{i=1} \ [\beta Z^G_i + (1-\beta)Z^N_i]}\ ,
\end{equation}
where $Z^S$, $Z^G_i$ and $Z^N_i$ are the Bayesian evidences (marginalised likelihoods) for $\mathcal{H}_S$, $\mathcal{H}_N$, and $\mathcal{H}_G$. $\alpha$ and $\beta$, are the prior odds for obtaining a signal $\alpha=P(\mathcal{H}_S)/P(\mathcal{H}_I)$ and the prior odds for obtaining a glitch $\beta=P(\mathcal{H}_G)/P(\mathcal{H}_I)$. \rs{In the case that these hypotheses are exhaustive and mutually exclusive, $\rho_{BCR}$ is the optimal discriminator between signals, glitches or noise.}

\rs{I suggest taking a pause here and talking about how we plan to use the BCR. You can recycle a lot of information from Isi+. A *key* point is that IMBH signals look a lot like glitches and we computing $\rho_{BCR}$ and we don't want to make a guess at a distribution for glitches and signals in this mass space (because we don't really know anything!) Looking at how the BCR is distributed for the background can give us some confidence that we construct a good discriminator between background and foreground
So, we don't take it at face value because we don't think we have a good model for the signal and/or glitch distribution. Thus, we "rank" the BCR of foreground triggers based on that of background triggers which allows us to compute a frequentist estimator for its significance. I suggest moving the current section II.E to be directly after this subsection. Note that you don't need to say where background triggers come from yet -- it suffices to say that background triggers are just data known to contain no coherent GW signals. }
As the rate of signal and glitches are unknown, these priors $\alpha$ and $\beta$ are tuned to maximise the \bcr distributions for background data (coherent signal-free data) and simulated signals~\cite{BCR1}.

\subsection{Bayesian Evidence Evaluation}
\subsubsection{Noise Model}
\rs{I suggest moving this subsection to an appendix}
We assume that each detector's noise is Gaussian and stationary over the period being analysed~\cite{ligo_psd}. In practice, we assume that the noise has a mean of zero that the noise variance $\sigma^2$ is proportional to the noise power spectral density (PSD) \psd of the data. Using the \psd, for each data segment $d_i$ in each of the $i$ detectors in a network of $D$ detectors, we can write 
\begin{equation}
\label{eq:zn}
Z^N_i = \mathcal{N}(d_i) = \frac{1}{2\pi \psd_i} \ \text{exp}\left(-\frac{1}{2} \frac{d_i}{\psd_i} \right) \ ,
\end{equation}
where $\mathcal{N}(d_i)$ is a normal distribution with $\mu=0$ and $\sigma^2\sim \psd$.

@@

\section{Bayesian Evidence Evaluation}
\subsection{Noise Model}
\rs{I suggest moving this subsection to an appendix}
We assume that each detector's noise is Gaussian and stationary over the period being analysed~\cite{ligo_psd}. In practice, we assume that the noise has a mean of zero that the noise variance $\sigma^2$ is proportional to the noise power spectral density (PSD) \psd of the data. Using the \psd, for each data segment $d_i$ in each of the $i$ detectors in a network of $D$ detectors, we can write 
\begin{equation}
\label{eq:zn}
Z^N_i = \mathcal{N}(d_i) = \frac{1}{2\pi \psd_i} \ \text{exp}\left(-\frac{1}{2} \frac{d_i}{\psd_i} \right) \ ,
\end{equation}
where $\mathcal{N}(d_i)$ is a normal distribution with $\mu=0$ and $\sigma^2\sim \psd$.

\subsection{Coherent Signal Model}
We model coherent signal using a binary black hole waveform template \template, where the vector \parameters contains a point in the 15 dimensional space describing precessing binary-black hole mergers. For the signal to be coherent, \parameters must be consistent in each 4-second data segment $d_i$ for a network of $D$ detectors, Hence, the coherent signal evidence is calculated as
\begin{equation}
\label{eq:zs}
Z^S = \int\limits_{\parameters} \prod\limits^{D}_{i=1} \left[ \mathcal{L}(d_i|\template)\right] \pi(\parameters | \mathcal{H}_S)\  \text{d}\parameters \ ,
\end{equation}
where $\pi(\parameters| \mathcal{H}_S)$ is the prior for the parameters in the coherent signal hypothesis, and $\mathcal{L}(d_i|\template))$ is the likelihood for the coherent signal hypothesis that depends on the gravitational wave template \template and its parameters \parameters.

\subsection{Incoherent Glitch Model}
Finally, as glitches are challenging to model and poorly understood, we follow \citet{bci} and utilise a surrogate model for glitches: the glitches are modelled using gravitational wave templates  \template with uncorrelated  parameters amongst the different detectors such that  $\parameters_i \neq \parameters_j$ for two detectors $i$ and $j$~\cite{bci}.  Modelling glitches with \template captures the worst case scenario: when glitches are identical to gravitational wave signals (excluding coherent signals). Thus, we can write $Z^G_i$ as 
\begin{equation}
\label{eq:zg}
Z^G_i = \int\limits_{\parameters} \mathcal{L}(d_i|\template)\ \pi(\parameters| \mathcal{H}_G)\  \text{d}\parameters  \ ,
\end{equation}
where $\pi(\theta| \mathcal{H}_G)$ is the prior for the parameters in the incoherent glitch hypothesis.

\section{Tuning the prior-odds}\label{sec:tuning-prior-odds}
